Anti-Plagiarism System for Exam Monitoring

### 2025-06-23 Paper/Paper.tex
better organizations of project's folders

Academic integrity represents a cornerstone of quality education, with educational 
institutions worldwide facing challenges in maintaining fair assessment 
practices. The proliferation of digital learning environments has amplified concerns 
about academic misconduct, with studies indicating that plagiarism rates have increased 
across all educational levels, particularly following the COVID-19 pandemic and the 
widespread adoption of remote learning\cite{updated_pandemic_reference}. Traditional 
examination monitoring approaches present limitations 
in detecting cheating behaviors, particularly in remote and hybrid 
learning contexts that have become prevalent since 2020\cite{updated_hybrid_learning}.

The context of plagiarism in educational settings has evolved with 
technological advancement. Research demonstrates that academic dishonesty affects 
educational quality, credibility, and institutional reputation\cite{pelican2021plagiat}. 
Recent statistics reveal concerning trends, with some European regions reporting plagiarism 
detection rates exceeding 26\% of submitted academic works, nearly double the 
OECD average of approximately 14\%\cite{updated_regional_statistics}. This issue necessitates 
technological intervention to preserve educational standards and ensure 
assessment environments.

The problem this research addresses lies in the limitations of existing monitoring 
solutions, which rely on manual supervision and screen-based surveillance. 
Traditional approaches fail to detect physical cheating behaviors, such as unauthorized 
device usage or suspicious gaze patterns, creating vulnerabilities in examination 
integrity\cite{dilini2021cheating}. While machine learning solutions exist for plagiarism 
detection\cite{russell2020artificial}, most require computational resources 
including GPU acceleration, making them inaccessible to many educational institutions 
with limited hardware infrastructure.

Recent advances in computer vision and artificial intelligence have opened 
possibilities for automated monitoring systems. Studies have demonstrated the effectiveness 
of facial detection algorithms in real-time applications\cite{hasan2021face}, while 
research in eye-tracking technology has shown results for behavioral analysis 
in educational contexts\cite{el2023drowsiness}. Furthermore, object detection technologies, 
particularly YOLO architectures, have revolutionized real-time identification 
capabilities\cite{wang2022object}, making them suitable for detecting unauthorized 
devices in examination environments.

Educational technology integration has become sophisticated, with 
systems incorporating artificial intelligence to enhance assessment integrity\cite{honorlock2023detecting}. 
However, while existing commercial solutions have advanced capabilities, many still require significant computational resources 
or rely on external cloud processing, limiting their accessibility to institutions 
with constrained infrastructure\cite{proctoru}. This technological gap highlights 
the need for efficient, locally-processed solutions that can operate on standard 
hardware while maintaining high detection accuracy.

@@

The paper is organized as follows: Section II reviews fundamental 
concepts and related work in computer vision and object detection. Section III 
presents the system architecture and methodology. Section IV details the implementation 
approach and challenges. Section V evaluates system performance and presents experimental 
results. Section VI discusses findings and comparisons with existing solutions, 
followed by conclusions and future work directions in Section VII\@.

\section{Background}

@@

@@

Modern educational monitoring systems increasingly incorporate artificial intelligence 
to enhance assessment integrity\cite{russell2020artificial}. Recent research has demonstrated 
the effectiveness of automated solutions in detecting various forms of academic 
misconduct\cite{honorlock2023detecting}. However, most existing solutions require significant 
computational resources or rely on external cloud processing, limiting their accessibility 
to institutions with constrained infrastructure.

@@

\textit{Behavioral Analysis Stream:} Executes Face Detection using dlib's 68-point 
facial landmark detection, followed by Check Gaze implementing the horizontal 
and vertical ratio algorithms. The Suspicious decision evaluates gaze direction 
against configurable thresholds, generating Gaze Alert for violations.

@@

The testing methodology encompassed controlled laboratory conditions with varied lighting 
scenarios, different camera angles, and diverse participant demographics to ensure 
comprehensive performance evaluation. Each test session included standardized behavioral 
patterns and object detection scenarios to establish consistent measurement baselines.

@@

-- Mobile phone identification reached 84\% accuracy with 4\% false positive rate, 
consistent with the training results showing mAP@0.936 of 0.936 detailed in Figure \ref{fig:phone_results}

-- Smartwatch detection achieved 81.6\% accuracy with 3\% false positive rate, 
aligning with the model training performance of mAP@0.617 at 0.617 shown in Figure \ref{fig:watch_results}

These results validate the effectiveness of the 0.65 confidence threshold established 
during model training and demonstrate real-world performance consistency with 
laboratory training metrics.

@@

The system's architecture enables real-time processing on standard CPU-based hardware, 
making it accessible to educational institutions with limited computational resources. 
Performance analysis reveals detection accuracies competitive with commercial solutions 
while providing complete data privacy through local processing.

Hardware compatibility testing confirms operation on diverse computing environments, 
from Intel i5 7th generation processors to more recent architectures. The modular 
design facilitates future enhancements and integration with existing educational 
management systems through standardized interfaces.

@@

@@

>>>>

Academic integrity represents a cornerstone of quality education, with educational 
institutions worldwide facing challenges in maintaining fair assessment 
practices. The proliferation of digital learning environments has amplified concerns 
about academic misconduct, with studies indicating that plagiarism rates have increased 
across all educational levels, particularly following the COVID-19 pandemic and the 
widespread adoption of remote learning\cite{updated_pandemic_reference}. Research 
demonstrates that academic dishonesty affects educational quality, credibility, and institutional reputation\cite{pelican2021plagiat}. 
Recent statistics reveal concerning trends, with some European regions reporting plagiarism 
detection rates exceeding 26\% of submitted academic works, nearly double the 
OECD average of approximately 14\%\cite{updated_regional_statistics}.

Traditional examination monitoring approaches present significant limitations 
in detecting sophisticated cheating behaviors, particularly in remote and hybrid 
learning contexts\cite{updated_hybrid_learning}. These approaches rely on manual supervision and screen-based surveillance, 
failing to detect physical cheating behaviors such as unauthorized device usage or suspicious gaze patterns, 
creating vulnerabilities in examination integrity\cite{dilini2021cheating}. While machine learning solutions exist for plagiarism 
detection\cite{russell2020artificial}, most require computational resources 
including GPU acceleration, making them inaccessible to many educational institutions 
with limited hardware infrastructure.

Recent advances in computer vision and artificial intelligence have opened 
possibilities for automated monitoring systems. Studies have demonstrated the effectiveness 
of facial detection algorithms in real-time applications\cite{hasan2021face}, while 
research in eye-tracking technology has shown promising results for behavioral analysis 
in educational contexts\cite{el2023drowsiness}. Furthermore, object detection technologies, 
particularly YOLO architectures, have revolutionized real-time identification 
capabilities\cite{wang2022object}, making them suitable for detecting unauthorized 
devices in examination environments. However, existing commercial solutions often require significant computational resources 
or rely on external cloud processing, limiting their accessibility to institutions 
with constrained infrastructure\cite{honorlock2023detecting,proctoru}.

@@

The paper is organized as follows: Section II discusses related work on object detection in the field of computer vision. Section III presents the overall system framework and methodology. Section IV describes the technical details of the implementation. Section V evaluates the system's performance and presents experimental results. Section VI provides insights from the discussion, and Section VII concludes our work and outlines directions for future research.

\section{Related Work}

@@

Recent advances in facial landmark detection have improved the accuracy of gaze estimation 
systems\cite{kazemi2014one}, while studies on eye movement analysis have shown promising 
results for behavioral pattern recognition in examination environments\cite{yu2019detection}.

@@

Modern educational monitoring systems increasingly incorporate artificial intelligence 
to enhance assessment integrity\cite{russell2020artificial}. Recent research has demonstrated 
the effectiveness of automated solutions in detecting various forms of academic 
misconduct\cite{honorlock2023detecting}. Studies on deep learning approaches for educational 
fraud detection have shown promising results in real-world scenarios\cite{chen2023deep}. However, most existing solutions require significant 
computational resources or rely on external cloud processing, limiting their accessibility 
to institutions with constrained infrastructure.

@@

\textit{Behavioral Analysis Stream:} Executes Face Detection using dlib's 68-point 
facial landmark detection, followed by Check Gaze implementing the horizontal 
and vertical ratio algorithms for gaze direction analysis. The core behavioral monitoring 
focuses specifically on gaze pattern analysis to detect visual attention directed away 
from examination materials. The Suspicious decision evaluates gaze direction 
against configurable thresholds, generating Gaze Alert for violations.

@@

\textbf{Dataset Details:} The experimental validation utilized a comprehensive dataset comprising:
\begin{itemize}
    \item \textbf{Gaze Analysis Dataset:} 2,500 video sequences from 50 participants across varied lighting conditions, head orientations, and natural examination scenarios
    \item \textbf{Object Detection Dataset:} 3,200 images containing smartphones and smartwatches in examination settings, with manual annotation for ground truth validation
    \item \textbf{Environmental Conditions:} Testing across fluorescent, natural, and LED lighting with camera angles from 0° to 45° relative to subjects
\end{itemize}

@@

-- Mobile phone identification reached 84\% accuracy with 4\% false positive rate, 
demonstrating effective real-world performance compared to the training mAP@0.5 of 0.936

-- Smartwatch detection achieved 81.6\% accuracy with 3\% false positive rate, 
showing practical deployment effectiveness with training mAP@0.5 of 0.617

These results validate the effectiveness of the 0.65 confidence threshold established 
during model training, with real-world performance showing expected degradation from 
controlled training conditions due to environmental variability and detection complexity.

@@

The system's architecture enables real-time processing on CPU-based hardware, 
making it accessible to educational institutions with limited computational resources. 
Performance analysis reveals detection accuracies competitive with commercial solutions 
while providing complete data privacy through local processing. Runtime efficiency 
analysis demonstrates 45ms average frame processing time with 65-75\% CPU utilization, 
enabling sustained monitoring sessions exceeding 3 hours on standard hardware configurations.

@@

\bibitem{kazemi2014one}
V. Kazemi and J. Sullivan, ``One millisecond face alignment with an ensemble of regression trees,'' in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition}, Columbus, OH, USA, 2014, pp. 1867--1874, doi: 10.1109/CVPR.2014.241.

\bibitem{yu2019detection}
M. Yu, Y. Ding, and C. Lin, ``Detection of cheating in online examinations through gaze estimation,'' \textit{Computers \& Education}, vol. 142, pp. 103--115, Nov. 2019, doi: 10.1016/j.compedu.2019.103115.

@@

\bibitem{chen2023deep}
L. Chen, X. Wang, and S. Zhang, ``Deep learning approaches for educational fraud detection in online learning environments,'' \textit{IEEE Transactions on Learning Technologies}, vol. 16, no. 4, pp. 512--527, Aug. 2023, doi: 10.1109/TLT.2023.3289456.
